---
title: "NBA pl_clean_f"
author: "Colin Busby"
date: '2022-04-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(randomForest)
library(parallel)
library(rpart)
library(readr)
library(janitor)
library(car)
library(MASS)
library(Hmisc)
```
# Multiple Linear Regression

## Look at salaries of players

Look at the salaries of players and determine the biggest factors in determining salary.

### Full Model

`Team`, `Position`, and `College` removed from model after being turned into factors `fTeam`, `fPosition`, and `College`, respectively.

`Points`, `Rebounds`, and `Assists` not included due to salaries being determined before beginning of season, and the stats are from the end of the season.

```{r}
lmfull <- lm(Salary ~ fTeam+fPosition+fCollege+Age+Height_i+Weight, data=pl_clean_f)
summary(lmfull)
```

### Removal of `fTeam` Due to lack of significance

```{r}
lm_r1 <- lm(Salary ~ fPosition+fCollege+Age+Height_i+Weight, data=pl_clean_f)
summary(lm_r1)
```

### Removal of `fCollege` Due to lack of significance

With several exceptions (`Weber State,Lehigh,Fresno State,Davidson`), the college

```{r}
lm_r2 <- lm(Salary ~ fPosition+Age+Height_i+Weight, data=pl_clean_f)
summary(lm_r2)
```

### Removal of `Height_i` Due to lack of significance

Given `Height_i`s p-value of `0.994822`, it is removed from the model.

```{r}
lm_r3 <- lm(Salary ~ fPosition+Age+Weight, data=pl_clean_f)
summary(lm_r3)
```



