---
title: "Multiple Linear Regression"
author: "Colin Busby"
date: '2022-04-16'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ~~USE `pl_clean_f` instead of `ap_clean_f`, as it contains `Points`, `Rebounds`, and `Assists`~~ Don't, it fucks up the dimensions(?), and I haven't figured out why or how to fix it

```{r}
library(tidyverse)
library(randomForest)
library(parallel)
library(rpart)
library(readr)
library(janitor)
library(car)
library(MASS)
library(Hmisc)
```

## Data Exploration

```{r}
dim(ap_clean_f)
```

```{r}
str(ap_clean_f)
```

```{r}
pairs(Salary ~ Age + Height_i + Weight, data = ap_clean_f)
```

## Model Creation

```{r}
summary(lmfull <- lm(Salary ~ fTeam+fPosition+fCollege+Age+Height_i+Weight, data=ap_clean_f))
```

```{r}
summary(lm_2 <- lm(Salary ~ fPosition+fTeam+Age+Height_i+Weight, data=ap_clean_f))
```

```{r}
summary(lm_3 <- lm(Salary ~ fPosition+Age+Height_i+Weight, data=ap_clean_f))
```

```{r}
summary(lm_3 <- lm(Salary ~ fPosition+Age+Weight, data=ap_clean_f))
```

```{r}
summary(lm_4 <- lm(Salary ~ fPosition+Age+Weight+fPosition:Age+fPosition:Weight+Age:Weight, data=ap_clean_f))
```

```{r}
anova(lm_3 ,lm_4, lm_2,lmfull)
```

## Check Model with AIC

```{r}
lm_step <- step(lmfull)
```

```{r}
anova(lm_4, lm_step, lm_2, lmfull)
```

## Model Checking

Right-hand outliers in `Normality of Residuals` are most likely the highly-paid players such as Stephen Curry (Need to check).

`Homogeneity of Variance` graph shows non-homogeneous variance, need to fix (or attempt to).

```{r}
performance::check_model(lm_step)
```

```{r}
performance::check_model(lm_4)
```

## Random Forest

```{r}
set.seed(999)
rf1 <- randomForest(Salary ~ fTeam+fPosition+Age+Height_i+Weight, data=ap_clean_f, importance=TRUE)
rf_preds <- predict(rf1, type = "response")
varImpPlot(rf1, type=1, main=" ")
```

```{r}
set.seed(999)
rf2 <- randomForest(Salary ~ fPosition+Age+Weight+fPosition:Age+fPosition:Weight+Age:Weight, data=ap_clean_f, importance=TRUE)
rf_preds_1 <- predict(rf2, type = "response")
varImpPlot(rf2, type=1, main=" ")
```

```{r}
set.seed(999)
rf3 <- randomForest(Salary ~ .-Team-Position-College-fCollege-Height-Name, data=ap_clean_f, n.var = 10, importance=TRUE)
rf_preds_3 <- predict(rf3, type = "response")
varImpPlot(rf3, type=1, main=" ")
```

## Plots

```{r}
ggplot(data=ap_clean_f) +
  geom_point(mapping = aes(x = Age, y = Salary)) + 
  facet_wrap(~ fPosition, nrow = 3)
```

```{r}
ggplot(data = ap_clean_f, mapping = aes(x = Age, y = Salary)) +
  geom_point(mapping = aes(color = fPosition)) +
  geom_smooth(data = filter(ap_clean_f, fPosition == "PG"), color = "red", se = FALSE)
```

## Random Forest Plots (Experimenting)

```{r}
#install.packages("plotmo")
library(plotmo)   # for the plotres function
plotres(lm_step)
```

```{r}
# Plots the trees of a randomForest

#install.packages("party")
library(party)
x <- ctree(Salary ~ fPosition + Age + Weight, data=ap_clean_f)
plot(x, type="simple")
```

```{r}
# `plot(randomForest(), log="y")` Plot the error rates or MSE of a randomForest object

plot(randomForest(Salary ~ fPosition + Age + Weight, data=ap_clean_f, keep.forest=FALSE, ntree=500), log="y")
```

## Questions

It would be interesting to see, barring super-stars like Steph Curry, how race factors into the salaries of NBA players. Do white players make more than Black players? Is it the opposite, or is it a non-factor?
