---
title: "NBA Salaries"
author: "Colin Busby & Hem Charan Bagul Krishnamurthy"
date: "`r Sys.Date()`"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Plan/Outline

Intro
- Question
- Data description
  - Variables
  - Dimensions/Structure
- Initial Data
  - Cleaning of Data
    - Why and How
  - Recoding for logistic regression
  - Graphs
    - Pairs()
    - ggpairs()
    - Histograms (skewness)
    - Shapiro and Breusch
    - Testing for Polnomials
- Model Building
  - Full(Initial) Model
  - Stepwise Model
  - Logistic
  - Random Forest

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

