---
title: "Salary MLR 2nd Draft"
author: "Colin Busby"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(randomForest)
library(parallel)
library(rpart)
library(readr)
library(janitor)
library(car)
library(MASS)
library(Hmisc)
library(rstatix)
library(forcats)
library(GGally)
library(performance)
```

# Intro

blahblahblah Hi

Question - How do the various stats from the ESPN website explain the salaries of NBA players?

Where data is from

Variables explained

## Data Exploration

```{r}
# Load dataset

players <- read_csv("players.csv")
active_p <- read_csv("active_players_2.csv")
```

```{r}
# Structure/Dimensions of the data.frame
str(active_p)
```

```{r}
# Number of columns containing `NA` or `NaN`
colSums(is.na(active_p))
```

## Data Cleaning

```{r}
#identifying the rows with NAs
rownames(active_p)[apply(active_p, 2, anyNA)]

#removing all `Salary` observations with NAs
ap_cl <- active_p[!(is.na(active_p$Salary)), ]
```

```{r}
# Checking for leftover `NA` after cleaning
rownames(active_p)[apply(ap_cl, 2, anyNA)]
```

```{r}
# Transforming Team,Position,College into factors
ap_cl_f <- transform (
  ap_cl,
  fTeam = as.factor(Team),
  fPosition = as.factor(Position),
  fCollege = as.factor(College)
)
```

```{r}
# Rename `fCollege` factor level "nan" as "None"
# levels(ap_cl_f$fCollege)[levels(ap_cl_f$fCollege)=="nan"] <- "None"
#levels(ap_cl_f$fCollege)
```

```{r}
str(ap_cl_f)
```

## Model Building

### Initial Graphs

```{r}
# The diagonal consists of the densities of the three variables and the upper panels consist of the correlation coefficients between the variables.

ggpairs(ap_cl_f, columns = c(9,4,6,7), cardinality_threshold = NULL)
```

### Full(Initial) Model

```{r}
lm1 <- lm(Salary ~ fTeam+fPosition+fCollege+Age+Height_i+Weight, data=ap_cl_f)
summary(lm1)
```


```{r}
summary(lm2 <- lm(Salary ~ fPosition+fCollege+Age+Height_i+Weight, data=ap_cl_f))
```

```{r}
summary(lm3 <- lm(Salary ~ fPosition+fCollege+Age+Weight, data=ap_cl_f))
```

```{r}
summary(lm4 <- lm(Salary ~ fPosition+Age+Weight, data=ap_cl_f))
```

```{r}
anova(lm4,lm3,lm2,lm1)
```
### Stepwise Model Selection

```{r}
summary(lm_step <- step(lm1))
```

## Model Checking

### Plots

```{r}
 par(mfrow=c(1,2), mar=c(4.5, 4.5, 2, 2))
 plot(lm1, 1:2)
```

```{r}
performance::check_model(lm_step)
```


### Outliers

While there are outliers in the data, their removal does not change the final model, in fact decreasing the adjusted $R^2$.  Therefore, they shall be left in the dataset.

```{r}
# Remove outliers from dataframe
ap_cl_f_wo <- ap_cl_f[-which(abs(rstandard(lm_step)) > 2),]
```

```{r}
lm1_wo <- lm(Salary ~ fTeam+fPosition+fCollege+Age+Height_i+Weight, data=ap_cl_f_wo)
summary(lm_step_wo <- step(lm1_wo))
```

### Box Cox

```{r}
boxcox(lm1,lambda=seq(-1, 1, by=0.05))
```

```{r}
summary(lm1_p1 <- powerTransform(lm1))
```

```{r}
Y1 <- coef(lm1_p1, round=TRUE)
summary(lm1_trans <- lm(bcPower(Salary, Y1)~fTeam + fPosition + fCollege + Age + Height_i + Weight, data = ap_cl_f))
```

```{r}
summary(lm_step_trans <- step(lm1_trans))
```

```{r}
performance::check_model(lm_step_trans)
```

```{r}
performance::compare_performance(lm1,lm2,lm3,lm_step, rank = TRUE)
```
```{r}
compare_performance(lm_step_trans,lm1_trans, rank = TRUE)
```

```{r}
plot(binned_residuals(lm_step))
```

